---
title: "Project 2: Improvements to Comparing the Statisitics of Countries in the World"
author: "Pixuan Shen"
output: html_document
date: "2024-11-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(janitor)
library(ggplot2)
library(maps)
library(patchwork)
```

Summary of changes and improvements: 

The first visualization created by my new code is very similar to Project 1, however, I changed it so that it will only highlight countries that the user tells the code to highlight. I put the map creation inside of a loop, so that the loop will cancel if the user gives a country that is either spelled incorrectly or does not exist. In the second visualization, the code and plot is completely different from that of Project 1. For the second visualization, I created a loop and put a conditional within the loop to generate a map of countries that fit a certain GDP per capita threshold. The loop will tell the user that their specified GDP per capita is too high and no countries fit the criteria if that is ever the case. The code has also been updated in both visualizations to fit the best practices criteria. The biggest thing that was done was editing the names and putting underscores to make the code easier to read and adding spaces wherever necessary.

```{r}
country_dataset <- read.csv("data/countries_of_the_world.csv")

country_dataset <- country_dataset %>%
  mutate(arable = as.numeric(gsub(",", ".", Arable)))

map_data_world <- map_data("world")

country_dataset <- country_dataset %>%
  mutate(Country = case_when(
    Country == "United States" ~ "usa",
    TRUE ~ Country
  ))

map_data_world <- map_data_world %>%
  mutate(region = tolower(region))

country_dataset <- country_dataset %>%
  mutate(Country = tolower(Country))

map_data_world <- left_join(map_data_world, country_dataset, by = c("region" = "Country"))

highlight_arable_land <- function(highlight_countries) {
  highlight_data <- map_data_world %>%
    filter(region %in% highlight_countries & !is.na(arable))
  
  if (nrow(highlight_data) > 0) {
    ggplot(map_data_world, aes(x = long, y = lat)) +
      geom_polygon(aes(group = group), fill = "white", color = "gray") +
      geom_polygon(data = highlight_data, aes(fill = arable, group = group), color = "gray") +
      scale_fill_gradient(low = "lightblue", high = "purple", na.value = "gray") +
      theme_minimal() +
      labs(
        fill = "% Arable Land",
        title = "Countries by Arable Land (%)",
        x = "Longitude",
        y = "Latitude"
      ) +
      coord_fixed()
  } else {
    stop("No data available for the specified countries.")
  }
}

highlight_arable_land(c("usa", "canada", "germany", "france", "brazil"))
```

```{r}
country_data <- read.csv("data/countries_of_the_world.csv")

country_data <- country_data %>%
  mutate(
    gdp_per_capita = as.numeric(gsub(",", "", GDP_per_capita)),
    country = tolower(case_when(
      Country == "United States" ~ "usa",
      TRUE ~ Country
    ))
  )

world_map <- map_data("world") %>%
  mutate(region = tolower(region))

map_data <- left_join(world_map, country_data, by = c("region" = "country"))

gdp_thresholds <- c(5000, 10000, 20000)

plots <- list()

for (threshold in gdp_thresholds) {
  map_data <- map_data %>%
    mutate(highlight = ifelse(gdp_per_capita > threshold, "Above Threshold", "Below Threshold"))
  
  if (any(map_data$highlight == "Above Threshold")) {
    p <- ggplot(map_data, aes(x = long, y = lat, group = group)) +
      geom_polygon(aes(fill = highlight), color = "gray") +
      scale_fill_manual(
        values = c("Above Threshold" = "purple", "Below Threshold" = "lightgray"), 
        na.value = "white"
      ) +
      labs(
        fill = "GDP per Capita",
        title = paste("Countries with GDP per Capita over", threshold, "USD"),
        x = "Longitude",
        y = "Latitude"
      ) +
      theme_minimal() +
      coord_fixed()
    
    plots[[as.character(threshold)]] <- p
  } else {
    stop(paste("No countries with GDP per capita above", threshold))
  }
}

combined_plot <- plots[[1]] + plots[[2]] + plots[[3]] + plot_layout(nrow = 3)

combined_plot
```


Dataset From Kaggle:

https://www.kaggle.com/datasets/fernandol/countries-of-the-world

References:

R Help Page - geom_smooth

R Help Page - Patchwork

https://stackoverflow.com/questions/61838700/query-on-how-to-make-world-heat-map-using-ggplot-in-r

https://stackoverflow.com/questions/34608767/add-geom-smooth-to-boxplot

https://stackoverflow.com/questions/12746667/increasing-the-line-thickness-of-geom-smooth

